### 2023-02-22 Manuscript simulations.Rmd
considering adding a simulation study using COMPAS

---
title: "Manuscript simulations"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{mdframed, caption}
  - \usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
  - \usepackage{float}
  - \floatplacement{figure}{H} 
  - \usepackage{longtable}
  - \usepackage{blkarray, bigstrut}
  - \usepackage{booktabs}
  - \usepackage{multirow}
---

\newpage

```{r setup, include = F}
rm(list = ls())

library(knitr)
hook_chunk <- knitr::knit_hooks$get('chunk')
knit_hooks$set(chunk = function(x, options) {

  # add latex commands if chunk option singlespacing is TRUE
  if(isTRUE(options$singlespacing)){
    return(sprintf("\\singlespacing\n %s \n\\doublespacing", hook_chunk(x, options)))
  } else{
    return(hook_chunk(x, options))
  }
})
knitr::opts_chunk$set(
  fig.align = "center",
  tidy = T,
  singlespacing = TRUE,
  cache = FALSE,
  fig.dim = c(10,8),
  message = FALSE,
  warning = FALSE,
  comment = NA,
  echo = F
)

# packages
packs <- c("dplyr", "nimble", "htmltools", "ggplot2", "sf", "Rcpp", "RcppArmadillo", "inline", "mvtnorm", "readr", "parallel", "xtable", "rstan", "coda", "vegan", "tidyr", "stringr", "tidyverse", "plotly", "CommEcol")
sapply(packs, require, character.only = T)
rm(packs)

# convenience
`%notin%` <- Negate("%in%")
options(mc.cores = parallel::detectCores())
```

```{r}
nimble_summary <- function(fit, warmup = nrow(fit[[1]])/2, thin = 1){
  # convert to coda for normal summary
  fit_warmup <- lapply(fit, function(x) x[(warmup+1):nrow(x),])
  coda_samples <- as.mcmc.list(lapply(fit_warmup, function(x) as.mcmc(
    x, start = warmup+1, end = nrow(fit), thin = thin
  )))

  sum <- summary(coda_samples)
  params <- dimnames(sum$statistics)[[1]]
  tmp_sum <- cbind(sum$statistics, sum$quantiles)

  # get r hat / n_eff
  mat <- matrix(NA, nrow = nrow(tmp_sum), ncol = 3)
  colnames(mat) <- c("Rhat", "ess_bulk", "ess_tail")
  for(i in 1:nrow(tmp_sum)){
    tmp <- sapply(fit, function(x) x[,i])
    mat[i,] <- c(Rhat(tmp), ess_bulk(tmp), ess_tail(tmp))
  }

  # out 
  out <- cbind(tmp_sum, mat)
  return(out)
}

fit_model <- function(seed = 1, code, data, constants, inits, niter, nchains, thin = 1){
  library(nimble)

  # R model
  model <- nimbleModel(code, constants, data)

  # C model
  model_c <- compileNimble(model)

  # R mcmc
  model_conf <- configureMCMC(model)
  model_conf$addMonitors(c("clus_id", "z"))

  # R mcmc
  mcmc <- buildMCMC(model_conf)

  # C mcmc
  mcmc_c <- compileNimble(mcmc, project = model_c)

  # run model
  out <- runMCMC(
    mcmc_c, 
    niter = niter, 
    nchains = nchains, 
    thin = thin, 
    init = inits,
    setSeed = seed
  )

  # out
  return(out)
}

ord_ls <- function(mcmc, d = 2, seed = NULL, force_K = NULL){
  # optional seed
  if(!is.null(seed)) set.seed(seed)

  # housekeeping
  mu <- lapply(mcmc, function(x) x[,which(grepl("mu", colnames(x)))])
  clus_id <- lapply(mcmc, function(x) x[,which(grepl("clus_id", colnames(x)))])
  other_ndx <- c(1:length(colnames(mcmc[[1]])))[-c(
    sort(c(
      which(grepl("clus_id", colnames(mcmc[[1]]))), 
      which(grepl("mu", colnames(mcmc[[1]])))
    ))
  )]
  other <- lapply(mcmc, function(x) x[,other_ndx])

  mu_mcmc <- do.call("rbind", mu)
  clus_mcmc <- do.call("rbind", clus_id)
  other_mcmc <- do.call("rbind", other)

  # helper function
  getmode <- function(v) {
    uniqv <- unique(v)
    uniqv[which.max(tabulate(match(v, uniqv)))]
  }

  # step 1 - determine number of non-empty cluster for each iteration
  n_clus_mcmc <- apply(clus_mcmc, 1, function(x) length(unique(x)))

  # step 2 - estimate the mode
  if(is.null(force_K)){
    n_clus_mode <- getmode(n_clus_mcmc)
  } else{
    n_clus_mode <- force_K
  }

  # step 3 - filter to iterations with n_clus_mcmc = n_clus_mode
  ## filter first
  which_equal_mode_index <- which(n_clus_mcmc == n_clus_mode)
  mu_mcmc_filtered <- mu_mcmc[which_equal_mode_index, ]
  clus_mcmc_filtered <- clus_mcmc[which_equal_mode_index, ]
  other_mcmc_filtered <- other_mcmc[which_equal_mode_index, ]

  ## remove empty clusters - loop
  mu_reduced <- matrix(NA, nrow = nrow(mu_mcmc_filtered), ncol = n_clus_mode * d)
  clus_reduced <- array(NA, dim = dim(clus_mcmc_filtered))
  for(i in 1:length(which_equal_mode_index)){
    clus_id_i <- sort(unique(clus_mcmc_filtered[i, ]))
    col_ndx <- which(
      stringr::str_sub(
        colnames(mu_mcmc_filtered), 
        start = stringr::str_locate(colnames(mu_mcmc_filtered), pattern = "\\[")[,1] + 1,
        end = stringr::str_locate(colnames(mu_mcmc_filtered), pattern = "\\,")[,1] - 1
      ) %in% clus_id_i
    )

    mu_reduced[i, ] <- mu_mcmc_filtered[i,col_ndx]
    colnames(mu_reduced) <- paste0("mu[", rep(1:n_clus_mode, d), ", ", rep(1:d, each = n_clus_mode), "]")
    clus_reduced[i, ] <- as.numeric(factor(clus_mcmc_filtered[i,]))
    colnames(clus_reduced) <- colnames(clus_mcmc_filtered)
  }

  # step 4 - create data matrix, cluster with k-means
  data_matrix <- matrix(
    c(mu_reduced), 
    nrow = nrow(mu_reduced) * n_clus_mode,
    ncol = d
  )
  k_means <- kmeans(data_matrix, centers = n_clus_mode)
  rho <- matrix(k_means$cluster, nrow = nrow(mu_reduced), ncol = n_clus_mode)

  # step 5 - check whether rho_i is a permutation of 1:n_clus_mode
  keep_ndx <- which(apply(rho, 1, function(x) all(sort(x) == 1:n_clus_mode)))
  mu_reduced <- mu_reduced[keep_ndx,]
  clus_reduced <- clus_reduced[keep_ndx,]
  other_reduced <- other_mcmc_filtered[keep_ndx, ]
  rho <- rho[keep_ndx,] %>% as.matrix(., ncol = n_clus_mode)

  # step 6 - relabel according to rho
  ## relabel mus first - works, but slow
  mu_relabeled <- array(NA, dim = dim(mu_reduced))
  colnames(mu_relabeled) <- colnames(mu_reduced)
  for(i in 1:nrow(mu_relabeled)){
    for(j in 1:ncol(mu_relabeled)){
      # get group id
      group_id <- stringr::str_sub(
        colnames(mu_reduced)[j], 
        start = stringr::str_locate(colnames(mu_reduced)[j], pattern = "\\[")[,1] + 1,
        end = stringr::str_locate(colnames(mu_reduced)[j], pattern = "\\,")[,1] - 1
      ) %>% as.numeric()

      dim_id <- stringr::str_sub(
        colnames(mu_reduced)[j], 
        start = stringr::str_locate(colnames(mu_reduced)[j], pattern = "\\,")[,1] + 1,
        end = stringr::str_locate(colnames(mu_reduced)[j], pattern = "\\]")[,1] - 1
      ) %>% as.numeric()

      # replace with proper value according to rho
      rho_i <- rho[i, ]
      mu_relabeled[i,j] <- mu_reduced[i, which(
        colnames(mu_reduced) == paste0("mu[", which(rho_i == group_id), ", ", dim_id, "]")
      )]
    }
  }

  ## relabel clus_id's
  clus_relabeled <- array(NA, dim = dim(clus_reduced))
  colnames(clus_relabeled) <- colnames(clus_reduced)
  for(i in 1:nrow(clus_relabeled)){
    rho_i <- rho[i,]
    clus_id_i <- clus_reduced[i,]
    clus_relabeled[i,] <- sapply(clus_id_i, function(x) rho_i[x])
  }

  out <- list(
    mu = mu_relabeled, 
    clus_id = clus_relabeled, 
    other = other_reduced,
    which_iter = which_equal_mode_index[keep_ndx]
  )

  return(out)
}

waic_dpord <- function(Y, mcmc, frame_ind, hier = F){
  if(hier){
    # get params 
    mu <- samples_relabeled$mu
    clus_id <- samples_relabeled$clus_id
    other <- samples_relabeled$other
    alpha <- other[,grepl("alpha", colnames(other))]
    beta <- other[,grepl("beta", colnames(other))]
    gamma <- other[,grepl("gamma[[]", colnames(other))]
    z <- other[,grepl("z", colnames(other))]
    theta <- other[,grepl("theta", colnames(other))]

    # get dimension
    d <- sapply(colnames(mu), function(x){
      str_sub(x, start = nchar(x)-1, end = nchar(x)-1)
    }) %>% as.numeric %>% max

    # compute lppd
    ## lik = alpha[i] + beta[j] + z[i] theta[j] + gamma[i]
    log_py <- matrix(NA, nrow = nrow(z), ncol = length(c(Y)))
    message("Calculating log likelihood")
    pb <- txtProgressBar(min = 0, max = nrow(log_py), style = 3, char = "=")
    for(iter in 1:nrow(log_py)){
      alpha_vec_s <- alpha[iter,]
      beta_vec_s <- beta[iter,]
      z_matrix_s <- matrix(z[iter,], ncol = d)
      z_matrix_s <- z_matrix_s[rep(1:nrow(z_matrix_s), table(frame_ind)),]
      theta_matrix_s <- matrix(theta[iter,], ncol = d)
      gamma_s <- gamma[iter, ]

      linpred <- rep(rep(alpha_vec_s, table(frame_ind)), ncol(Y)) +
        rep(beta_vec_s, each = nrow(Y)) +
        c(z_matrix_s %*% t(theta_matrix_s)) +
        rep(gamma_s, ncol(Y))
      pi <- unname(exp(linpred) / (1 + exp(linpred)))

      # fix numerical issues
      pi[which(pi > 0.99999)] <- 0.99999
      pi[which(pi < 0.00001)] <- 0.00001

      log_py[iter,] <- c(Y) * log(pi) + (1 - c(Y)) * log(1 - pi)
      setTxtProgressBar(pb, iter)
    }
    close(pb)

    # compute lppd
    message("Computing lppd")
    lppd <- sum(log(colMeans(exp(log_py))))

    # compute penalty
    message("Computing penalty")
    pwaic2 <- sum(apply(log_py, 2, function(x) var(x)))

    # compute waic
    waic <- -2*(lppd - pwaic2)
    out <- waic
  } else{
    # get params 
    mu <- samples_relabeled$mu
    clus_id <- samples_relabeled$clus_id
    other <- samples_relabeled$other
    alpha <- other[,grepl("alpha", colnames(other))]
    beta <- other[,grepl("beta", colnames(other))]
    z <- other[,grepl("z", colnames(other))]
    theta <- other[,grepl("theta", colnames(other))]

    # get dimension
    d <- sapply(colnames(mu), function(x){
      str_sub(x, start = nchar(x)-1, end = nchar(x)-1)
    }) %>% as.numeric %>% max

    # compute lppd
    ## lik = alpha[i] + beta[j] + z[i] theta[j] + gamma[i]
    log_py <- matrix(NA, nrow = nrow(z), ncol = length(c(Y)))
    message("Calculating log likelihood")
    pb <- txtProgressBar(min = 0, max = nrow(log_py), style = 3, char = "=")
    for(iter in 1:nrow(log_py)){
      alpha_vec_s <- alpha[iter,]
      beta_vec_s <- beta[iter,]
      z_matrix_s <- matrix(z[iter,], ncol = d)
      theta_matrix_s <- matrix(theta[iter,], ncol = d)

      linpred <- rep(alpha_vec_s, ncol(Y)) +
        rep(beta_vec_s, each = nrow(Y)) +
        c(z_matrix_s %*% t(theta_matrix_s)) 
      pi <- unname(exp(linpred) / (1 + exp(linpred)))

      # fix numerical issues
      pi[which(pi > 0.99999)] <- 0.99999
      pi[which(pi < 0.00001)] <- 0.00001

      log_py[iter,] <- c(Y) * log(pi) + (1 - c(Y)) * log(1 - pi)
      setTxtProgressBar(pb, iter)
    }
    close(pb)

    # compute lppd
    message("Computing lppd")
    lppd <- sum(log(colMeans(exp(log_py))))

    # compute penalty
    message("Computing penalty")
    pwaic2 <- sum(apply(log_py, 2, function(x) var(x)))

    # compute waic
    waic <- -2*(lppd - pwaic2)
    out <- waic
  }

  return(out)
}

getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

# Introduction

Simulations using COMPAS, similar to Hui.

# Compas example

Generates abundances, switch to presence-absence.

```{r}
set.seed(02212023)
coords <- rbind(
  mvtnorm::rmvnorm(15, c(30, 60), 30*diag(2)),
  mvtnorm::rmvnorm(15, c(60, 40), 30*diag(2)),
  mvtnorm::rmvnorm(15, c(80, 60), 30*diag(2))
)
plot(coords, col = rep(1:3, each = 15), main = "True latent coordinates")
mat <- compas(
  S = 30, 
  dims = 2,
  am = c(5, 5),
  beta.R = c(1, 1), 
  coords = coords, 
  n.quanti = 1,
  n.quali = 0.05, 
  add1 = 0.05
)
# mat[mat > 1] <- 1
mat <- mat[which(rowSums(mat) != 0),]
```

## NMDS

```{r}
nmds <- vegan::metaMDS(
  mat,
  distance = "bray", 
  k = 3
)
plot(nmds$points)
```

## DPORD

```{r}
library(gmp)
# probability mass function from Antoniak (1974)
antoniak <- function(k, alpha, n){
  alpha <- gmp::as.bigq(as.character(MASS::fractions(alpha)))

  # A.n first
  A.n <- function(x, n){
    sum(abs(gmp::Stirling1.all(n)) * x^(1:n))
  }

  # compute mass
  out <- as.numeric((abs(gmp::Stirling1(n, k)) * alpha^k) / A.n(alpha, n))
  return(out)
}
plot(sapply(1:10, function(x) antoniak(x, 1/1, 45)), main = "Prior mass on number of groups for n = 45, alpha = 1/2")
```

```{r}
# model based 
code <- nimbleCode({
  # priors
  phi ~ dunif(0, 10)
  ## site effects
  for(site in 1:nsites){
    alpha[site] ~ dnorm(0, 1)
  }

  ## species effects
  for(species in 1:nspecies){
    beta[species] ~ dnorm(0, 1)
  }

  ## z priors
  ### Dirichlet process mixture parameters
  clus_id[1:nsites] ~ dCRP(dp_con, size = nsites)
  dp_con ~ dgamma(1, 1)

  ### table parameters - fix covariance as identity
  for(i in 1:max_clus){
    mu[i, 1:d] ~ dmnorm(mu0[1:d], Lambda0[1:d, 1:d])
  }

  for(site in 1:nsites){
    # identity matrix for constraint
    z[site, 1:d] ~ dmnorm(mu[clus_id[site], 1:d], cov = S[1:d, 1:d])
  }

  # theta prior
  ## upper triangle = 0
  for(row in 1:(d-1)){
    for(col in (row+1):d){
      theta[row, col] <- 0
    }
  }

  ## diag > 0
  for(diag_element in 1:d){
    theta[diag_element, diag_element] ~ T(dnorm(0, sd = 1), 0, Inf)
  }

  ## lower diag of first d rows
  for(row in 2:d){
    for(col in 1:(row-1)){
      theta[row, col] ~ dnorm(0, sd = 1)
    }
  }

  ## all other elements
  for(row in (d+1):nspecies){
    for(col in 1:d){
      theta[row, col] ~ dnorm(0, sd = 1)
    }
  }

  # likelihood 
  for(site in 1:nsites){
    for(species in 1:nspecies){
      log(lambda[site, species]) <- alpha[site] + beta[species] + inprod(z[site,1:d], theta[species, 1:d])
      nbp[site, species] <- phi / (phi + lambda[site, species])
      Y[site, species] ~ dnegbin(prob = nbp[site, species], size = phi)
    }
  }
})
init_func <- function(){
  d <- 2
  nrow <- 40
  nspecies <- 11
  max_clus <- 40

  theta_init <- matrix(rnorm(d * nspecies), nspecies, d)
  diag(theta_init) <- abs(rnorm(d))
  theta_init[upper.tri(theta_init)] <- 0

  list(
    alpha = rnorm(nrow),
    beta = rnorm(nspecies),
    z = matrix(rnorm(nrow * d), nrow, d),
    theta = theta_init,
    dp_con = 1,
    clus_id = sample(c(1, 2), size = nrow, replace = T),
    mu = matrix(rnorm(d*max_clus), max_clus, d),
    phi = 1
  )
}

this_cluster <- makeCluster(3)
fit <- parLapply(
  cl = this_cluster,
  X = 1001:1003,
  fun = fit_model,
  code = code,
  data = list(
    Y = mat,
    S = diag(2),
    mu0 = rep(0, 2),
    Lambda0 = diag(2)
  ),
  constants = list(
    nsites = nrow(mat),
    nspecies = ncol(mat),
    d = 2,
    max_clus = 40
  ),
  inits = init_func,
  niter = 10000,
  nchains = 1,
  thin = 1
)
stopCluster(this_cluster)
```

```{r}
fit_ <- fit
fit <- lapply(fit, function(x) x[(nrow(x)/2 + 1):nrow(x), ])
samples_relabeled <- ord_ls(mcmc = fit[c(1:3)], d = 2, seed = 02212023)

z_mcmc <- samples_relabeled$other[,which(grepl("z", colnames(samples_relabeled$other)))]
z_tbl <- left_join(
  tibble(
    trace = c(z_mcmc),
    param = rep(colnames(z_mcmc), each = nrow(z_mcmc)),
    iter = rep(1:nrow(z_mcmc), length(colnames(z_mcmc))),
    obs_id = stringr::str_sub(
      param, 
      start = stringr::str_locate(param, pattern = "\\[")[,1] + 1,
      end = stringr::str_locate(param, pattern = "\\,")[,1] - 1
    ) %>% as.numeric(),
    dim = stringr::str_sub(
      param, 
      start = stringr::str_locate(param, pattern = "\\,")[,1] + 1,
      end = stringr::str_locate(param, pattern = "\\]")[,1] - 1
    ) %>% as.numeric()
  ),
  tibble(
    clus_id = c(samples_relabeled$clus_id),
    iter = rep(1:nrow(samples_relabeled$clus_id), length(colnames(samples_relabeled$clus_id))),
    obs_id = rep(1:ncol(samples_relabeled$clus_id), each = nrow(samples_relabeled$clus_id))
  ),
  by = c("iter", "obs_id")
)

clus_id <- lapply(fit, function(x) x[,which(grepl("clus_id", colnames(x)))])
clus_id_all <- do.call("rbind", clus_id)
ngroups_mcmc <- apply(clus_id_all, 1, function(x) length(unique(x)))
round(table(ngroups_mcmc) / length(ngroups_mcmc), 3)

z_tbl %>%
  group_by(obs_id, dim) %>%
  summarize(
    mean = mean(trace),
    clus = getmode(clus_id),
    clus_prob = mean(clus_id == clus)
  ) %>%
  pivot_wider(
    names_from = dim,
    values_from = mean, 
    names_prefix = "z"
  ) %>%
  ungroup() %>%
  mutate(
    cluster = factor(clus)
  ) %>% 
  mutate(
    obs_id = ifelse(obs_id > 7, obs_id + 1, obs_id)
  ) %>%
  ggplot() +
  geom_text(aes(x = z1, y = z2, col = cluster, label = obs_id)) +
  theme_bw() +
  # guides(col = "none") +
  labs(x = bquote(z[1]), y = bquote(z[2]))
```

# Simulation

How complex?